\Large{Adversarial Training against Systematic Uncertainty}

### 2016-09-29 tex/paper.tex
save

Nice and cozy abstract goes here...

@@

The discovery of new particles or physical phenomena in high energy physics
experiments, like those at the Large Hadron Collider, requires the observation
of statistically significant deviations from the predictions of the Standard
Model, the current model of the known fundamental particles and their
interactions.  Typically, discovery requires satisfying the \textit{5-sigma
rule}, whereby the deviation must be at least five standard deviations from
predictions, i.e. having a p-value of $p \lesssim 3 \times 10^{-7}$. The
evaluation of the statistical significance must not only account for statistical
uncertainties due to random fluctuations, but also for
systematic uncertainties that represent our incomplete knowledge of the theory
of particle interactions or of the properties of the experimental detection
apparatus. Systematic uncertainties can alter the expected rate of signal and
background classes, e.g. their priors, as well as the distributions of features.
As such, the signal class true-positive rate and the background class
false-positive rate of a classifier can change with variations in the systematic
uncertainty thereby decreasing the sensitivity of an analysis. In this paper, we
present a new strategy for training a classifier, based on adversarial
training~\citep{goodfellow2014generative}, which aims at building a classifier
whose decision function is invariant under variations due to systematic
uncertainties.  In this way, uncertainties on predictions from simulations will
have a reduced impact when computing the significance of potential deviations
after application of the discriminant model and thereby increase the sensitivity
of the analysis to signals from potential new physics sources. \glnote{Shall we
be more explicit somewhere and put a formula underlining the effect of having a
pivotal classifier on the resulting sensitivity? (versus a classifier whose
decision function depends on the nuisance.)}

>>>>

Systematic uncertainties in high energy physics, representing our incomplete
knowledge in the simulation of physical processes or in the detection apparatus,
may cause variations in the conditional distributions of features used for
training a classifier.  As such, the efficacy of a classifier may change under
the variations caused by systematic uncertainties.  In this paper, we repurpose
adversarial training as a means to produce a pivotal classifier which is
invariant to variations of the nuisance parameters which parameterize the
effects of systematic uncertainties. In the machine learning context, this
problem is closely related to those of domain adaptation and enforcing fairness
in classification. Following that line of work, the method proposed in this work
can be regarded as generalization that also supports the continuous case,
which can be viewed as handling infinitely many domains or as enforcing fairness
over continuous attributes.

@@

The discovery of new particles or physical phenomena in high energy physics
experiments, like those at the Large Hadron Collider, requires the observation
of statistically significant deviations from the predictions of the Standard
Model, the current model of the known fundamental particles and their
interactions.  Typically, discovery requires satisfying the \textit{5-sigma
rule}, whereby the deviation must be at least five standard deviations from
predictions, i.e. having a p-value of $p \lesssim 3 \times 10^{-7}$. The
evaluation of the statistical significance must not only account for statistical
uncertainties due to random fluctuations, but also for
systematic uncertainties that represent our incomplete knowledge of the theory
of particle interactions or of the properties of the experimental detection
apparatus. Systematic uncertainties can alter the expected rate of signal and
background classes, e.g. their priors, as well as the distributions of features.
As such, the signal class true-positive rate and the background class
false-positive rate of a classifier can change with variations in the systematic
uncertainty thereby decreasing the sensitivity of an analysis. In this paper, we
present a new strategy for training a classifier, based on adversarial
training~\citep{goodfellow2014generative}, which aims at building a classifier
whose decision function is invariant under variations due to systematic
uncertainties.  In this way, uncertainties on predictions from simulations will
have a reduced impact when computing the significance of potential deviations
after application of the discriminant model and thereby increase the sensitivity
of the analysis to signals from potential new physics sources.
